Network Structure, Efficiency, and Performance in WikiProjects

### 2018-04-06 paper-content.tex
Cleans up formatting, grammar, clarity.

\begin{abstract}
The internet has enabled collaborations at a scale never before possible,
but the best practices for organizing such large collaborations are still not clear.
Wikipedia is a visible and successful example of such a collaboration which might offer
insight into what makes large-scale, decentralized collaborations successful.
We analyze the relationship between the structural properties of WikiProject coeditor networks
and the performance and efficiency of those networks.
We confirm the existence of an overall performance-efficiency trade-off,
while observing that some projects are higher than others in both performance
and efficiency,
suggesting the existence factors correlating positively with both.
Namely, we find an association between low-degree coeditor networks
and both high performance and high efficiency.
We also confirm results seen in previous numerical and small-scale lab studies:
higher performance with less skewed node distributions,
and higher performance with shorter path lengths.
We use agent-based models to explore possible mechanisms for
degree-dependent performance and efficiency.
We present a novel local-majority learning strategy designed to satisfy properties
of real-world collaborations.
The local-majority strategy as well as a localized conformity-based strategy
both show degree-dependent performance and efficiency,
but in opposite directions,
suggesting that these factors depend on both network structure and learning strategy.
Our results suggest 
possible benefits to decentralized collaborations made of smaller,
more tightly-knit teams,
and that these benefits may be modulated by the particular learning strategies
in use.
\end{abstract}

@@

We look at WikiProjects on the English-language Wikipedia.
WikiProjects are collections of thematically related articles,
each with their own standards and norms.
When measuring the quality of collaborative projects,
there are at least two distinct measures to consider.
The first measure is short-term:
how effective a unit of work is at improving
the collaboration's output,
which we call {\em efficiency}.
The other measure is long-term:
the highest quality typically reached by an output,
which we call {\em performance}.
These two terms are often used interchangeably,
but we find it fruitful to distinguish between the two.
We find that Wikipedia exhibits an overall trade-off between performance and efficiency.
However, some WikiProjects are higher in both efficiency and performance
than others,
suggesting the existence of factors that correlate positively with both.

Our study focuses on the coeditor networks of each WikiProject:
which editors have edited at least one article in common?
These relationships represent possible flows of information.
We focus specifically on mean degree, degree skewness, and path length.
High-degree editors have more collaborators,
which can increase diversity and access to information at the possible
expense of higher coordination costs
\cite{hong_groups_2004,golub_naive_2010}.
Highly skewed degree distributions can amplify the biases of high-degree
editors while reducing the need for explicit coordination
\cite{kearns_experiments_2012}.
Networks with shorter path lengths allow information to travel more quickly
at the possible expense of less localized diversity
\cite{mason_propagation_2008,barkoczi_social_2016}.

In addition to our empirical study,
we use agent-based modeling to examine the consequences of specific
assumptions on networked collaboration.
We model individual behavior using a {\em social learning strategy} that
assumes agents 1. can only access a fraction of the model's state,
2. interact with others who share their concerns,
and 3. reconcile their preferences into a single state.
Our model is the first we are aware of to incorporate these assumptions,
which are present across many real-world collaborations
including Wikipedia.

Our main findings are:
\begin{itemize}
\item Despite an overall performance/efficiency trade-off,
WikiProjects having low-degree coeditor networks tend to have both higher performance and higher efficiency;
\item Short paths are associated with higher performance, consistent with a conformity-based learning strategy;
\item Structural inequality, as measured by degree skewness, is associated with lower performance;
\item Our agent-based model
shows that the efficiency and performance of collaborations can
depend on network degree,
and that the direction of that dependence varies with social learning strategy.
\end{itemize}

Our findings shed light on the importance of network structure for successful
collaboration.
These findings might be informative for future interventions which
recommend tasks based on how they will
influence network structure,
or interventions which seek to encourage the behavioral norms most
effective for existing network structure.

\section{Background and Related Work}
\label{sec:background}
The present paper investigates the relationship between social networks
and collaboration outcomes.
This connection has been explored by a number of theoretical,
numerical, and small-scale lab studies in the field of
{\em social learning}.
We expand the literature with a large-scale, empirical field study.
In much of the existing literature,
degree distribution correlates with outcome measures.
But aside from the naive Bayes case, it is unknown whether
the correlation is explained best by degree or by another structural property, such as characteristic path length.
In the empirical networks we study,
unlike artificial networks,
the structural properties vary independently,
making it easier to isolate individual
network properties that correlate with outcome variables.

{\bf Social Learning.} In {\em networked social learning},
agents are represented by nodes on a network
and can only interact with their neighbors.
Social learning tasks can be divided into cases where agents have {\em generated signals}
(independently noisy estimates of a true value)
and those where agents have {\em interpreted signals}
(solutions based on different selections of available data)
\cite{hong_interpreted_2009}.
The behavior of individual agents is described by their
{\em social learning strategy}.
For generated signals,
a naive Bayesian approach converges to the truth
when all agents have the same degree,
while the speed of convergence depends on the {\em spectral gap}
between the two largest eigenvalues of the network's adjacency matrix
\cite{degroot_reaching_1974,golub_naive_2010}.
Complex social learning tasks can also be modeled as the problem
of maximizing an objective function with many local maxima,
referred to as a {\em rugged landscape}
\cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,grim_scientific_2013,barkoczi_social_2016}.
Numerical simulations have shown that efficient networks
(those with short paths between nodes)
can result in faster convergence at the cost of a less optimal solution,
due to less time for exploration
\cite{mason_propagation_2008,grim_scientific_2013}.
However, when conformity-based social learning strategies are used,
efficient networks can sometimes find more optimal solutions than
inefficient ones \cite{barkoczi_social_2016}.

{\bf Lab experiments.} Lab-based experiments on networked collaboration
suggest a complex interaction between network topology and other factors.
While groups of networked human subjects perform very well on
difficult graph-coloring tasks, the best performing network architectures
(e.g., fully-connected vs. small-world) vary
from task to task \cite{kearns_experiments_2012}.
The same studies found that while human subjects tend to perform well on
some networks, they perform worst on self-organized
networks.
Similarly, some network topologies are able to reach faster decisions in the
presence of more information, while others show the opposite effect
\cite{kearns_experimental_2006}.
Based on lab experiments, Fowler and Christakis \shortcite{fowler_cooperative_2010}
suggest that individual decisions towards altruism are conditional on their
neighbor's behavior and ``contagious'' up to three degrees away.
Later experiments by Suri and Watts \shortcite{suri_cooperation_2011} confirmed the
existence of conditional altruism,
but concluded that altruistic
behavior only influences first-degree neighbors.

{\bf Digital Communities.}
Research on digital communities has also examined the role of 
diversity and inequality
in collaborative work and decision-making.
Using an agent-based model, Hong and Page \shortcite{hong_groups_2004} found that
diverse groups can outperform groups composed of the best individual
problem-solvers.
In sociology, research has focused on the relationship between
network structure and social capital.
Powerful individuals are often ``brokers''
who act as exclusive intermediaries between disconnected portions of the
social network \cite{silverman_patronage_1965}.
Similarly, successful innovation in organizations often occurs in ``structural
holes'' between groups \cite{granovetter_strength_1973}.
There is some evidence that groups with higher structural inequality
(degree skewness)
perform worse \cite{kearns_experiments_2012}.

@@

Across the broad range of work discussed above,
a few key themes emerge.
The efficiency and performance of collaborations are both important considerations
and vary depending on both network structure and type of task
\cite{kearns_experiments_2012}.
While generated signal models of social learning predict no relationship
between the two
\cite{golub_naive_2010},
contagion-style innovation models predict a trade-off
\cite{mason_collaborative_2012,barkoczi_social_2016}.
Such a trade-off has been observed in simulations and lab experiments on
collaboration
\cite{kearns_experiments_2012,grim_scientific_2013}.

@@

Many articles on Wikipedia belong to one or more WikiProject.
WikiProjects are groups of thematically-related articles
(e.g., articles related to Philosophy).
Information about the an article's associated WikiProjects
can be viewed on that
article's talk page (Figure \ref{fig:knitting}).
Each WikiProject has its own page and talk page,
containing information about community norms within the project
as well as discussions about individual articles.
WikiProjects are thus distinct communities, with distinct norms and processes.
These communities are the fundamental units of analysis in this paper.

@@

In addition to the above assessment classes, articles on Wikipedia can be tagged as
"good article" (GA) or "featured article" (FA) quality.
FA and GA determinations are made using a Wikipedia-wide consensus,
in parallel to any WikiProject-based evaluations.
FA articles are ``the best articles Wikipedia has to offer''
\cite{wikipedia_contributors_wikipedia:featured_2018}.
GA articles meet ``a core set of editorial standards`` but are ``not featured article quality''
\cite{wikipedia_contributors_wikipedia:good_2017}.
When an article is assigned GA or FA status,
WikiProject quality assessments are often updated to reflect that status.
For example, the article {\em Mewtwo} was assessed as GA status on October 5,
2009 and shortly afterwards
its quality assessment was changed from B to GA within both
{\em WikiProject Pok\'emon} and
{\em WikiProject Video Games}.
This example also illustrates a quirk of Wikipedia norms:
very often, articles pass to GA or FA directly from B, skipping A.
The majority of WikiProjects rarely use the A class quality assessment.

@@

For a WikiProject, efficiency quantifies how much participants can raise the
assessed quality of an article for a fixed amount of work.
We measure work by the number of revisions made.
Quality assessments are made through consensus of the project participants themselves,
so different projects can have different standards and practices for assessing article quality.
So the efficiency is not a measure of how quickly some objective measure of quality improves,
but rather of how quickly the project participants can reach consensus on the improvements that
need to be made and make those improvements.
Because our definition relies on assessment transitions,
we must define efficiency variables for
each of the project-level quality assessments: A, B, and C.
For a particular grade $G$,
we desire our definition of efficiency to meet the following conditions:
\begin{itemize}
\item{Strictly increasing in the number of articles reaching grade $G$ (with revision count fixed);}
\item{Strictly decreasing in the number of revisions (with transition count fixed);}
\item{Independent of WikiProject size: not affected by adding an article with the same efficiency.}
\end{itemize}

We now define an efficiency measure which meets the above criteria.
Let $T(W,G)$ be the set of article assessment transitions from below grade $G$
to grade $G$ or higher in project $W$.
Let $N(W,G)$ be the number of articles in project $W$ which ever transition
from below grade $G$ to grade $G$ (or higher).
Let $r(t)$ be the number of article revisions since its previous grade transition,
and let $g(t)$ be the number of grade levels crossed by transition $t$.
We quantify the efficiency $E(W,G)$ as the inverse of the mean number of revisions
per transition:
\beq
E(W,G)
&=&
\left[
\frac{1}{
N(W,G)
}
\sum_{t \in T(W,G)} \frac{r(t)}{g(t)}
\right]^{-1},
\eeq
Where the $g(t)$ term accounts for assessments raising article quality by several
grades,
dividing the revisions evenly between all grade levels achieved.

For performance, we wish to quantify how good articles tend to be when they reach a stable state.
Measuring performance is difficult for several reasons:
there is no objective measure of article quality available,
and articles are always changing, making it difficult to know which articles should be considered
complete or stable.
We use an extremely simple performance measure that gives surprisingly consistent results.
In addition to per-project quality assessment, articles can be given ``featured article'' or
``good article'' status.
The criteria for these statuses are consistent across all of Wikipedia,
and any editor can participate in the discussion and decision to award good or featured
status.
In other words, the good and featured statuses are more objective than per-project assessments.

@@

We would like to determine how the social network structure of
Wikipedia--the pattern of who interacts with whom--relates to
efficiency and performance.
There are several types of interactions we could focus on,
including:
coediting, user talk messages, and talk page replies.
We choose to focus on coediting: when two editors have made changes to the
same article or talk page.
While editors can communicate directly through user talk messages,
the number of such messages is small compared to the number of edits to article
and talk pages.
We also could have considered direct replies between editors on article talk
pages, but these replies are typically seen (and intended to be seen)
by everyone reading the talk page,
and are part of larger conversations.
When an editor views a page,
they are potentially viewing content from and interactions
between all editors who came before them,
motivating our choice to focus on the social network structure of
coeditors.

The {\em coeditor network} of a WikiProject consists of nodes representing editors,
and edges connecting any editors who have edited the same article.
The edges are directed, with the direction representing
{\em plausible information flow};
an edge from Alice to Bob exists if Alice edited an article and then Bob edited the same article at
a later time. Note that edges can exist in both directions. 
We make the simplifying assumption of unit weight for all edges.
We focus on three structural properties:
degree, characteristic path length, and min-cut.
Degree and characteristic path length have been shown to correlate with
performance and efficiency in some social learning settings
\cite{golub_naive_2010,mason_propagation_2008,grim_scientific_2013},
while min-cut can be interpreted as a measure of decentralization,
common feature of peer-produced communities such as Wikipedia
\cite{benkler_wealth_2006}.

The degree distribution is the simplest network property we analyze.
The in-degree (out-degree) of a node is the number of edges to (from) that node.
Taking the average of either in-degree or out-degree gives the same value:
the {\em mean degree} of the network.
In our context, the mean degree represents how many others a given editor has collaborated with.
We also consider the {\em skewness} of the in-degree and out-degree distributions.
A large positive degree skewness for a WikiProject coeditor network
implies that a small number of editors have a very large number of collaborators,
while a small positive value implies that the editors having the most collaborators
don't have many more than a typical editor.

@@

Our final network measure quantifies the connectivity of a project's coeditor network using
min-cut size.
The minimum $st$-cut between nodes $s$ and $t$ is the set of edges that must be removed in order that
no path exists from $s$ to $t$.
The minimum cut (min-cut) of a graph is the smallest minimum $st$-cut over all node pairs $st$. 
The size of the graph min-cut quantifies the connectivity of a graph,
but only incorporates information about edges lying on paths crossing the min-cut.
Instead, we use the mean size of all minimum $st$-cuts, which we refer to as the
{\em mean min-cut}.
This measure quantifies the number of redundant paths information can take through the network.
Networks with higher redundancy are more resilient to errors on one path \cite{albert_error_2000}
and allow innovations to propagate through complex contagion,
in which innovations are only adopted after multiple exposures \cite{centola_complex_2007}.

The mean path and min-cut are computationally intensive,
requiring distance and minimum $st$-cut calculations for each all node pairs.
For larger projects, these calculations are impractical and we thus employed sampling to determine mean path length and mean
min-cut.
For mean path length, source nodes were sampled, and path length was calculated to all destination nodes
from each of these.
For min-cut, node pairs were sampled.
In both cases, stratification was used to ensure the same number of nodes were were sampled from each of
12 node degree quantiles.
We estimated the error due to sampling by determining true values for a medium-sized project,
and calculating error as a function of sample-size.
Sample sizes were chosen such that relative error was below 10\%.
Even with sampling, however, it was impractical to calculate these properties for the largest projects,
so we exclude the 183 largest projects from the analysis.

@@

We also find that mean min-cut is highly correlated with degree ($r=0.980$, $p<0.001$),
so we exclude min-cut from regression models to prevent collinearity.
The high correlation between mean degree and min-cut implies that in most cases
the minimum $st$-cut is simply the either set of edges from $s$ or the set of edges to $t$.
The rarity of non-trivial min-cuts suggests that WikiProject coeditor networks have very few central
bottlenecks and are thus highly decentralized.

@@

To study the relationship between network structure, efficiency, and performance,
we model the performance and efficiency of WikiProjects using ordinary least-squares linear regression.
Each WikiProject is a single observation.
The models include each project's coeditor network properties as independent variables.
We also include the following project-level variables to control for confounding factors.
\begin{description}
\item[C-efficiency]
(Performance only.)
Quantifies how quickly a WikiProject improves articles.
Efficiencies for different grades are highly correlated,
so we include only one.
\item[Connected fraction]
Fraction of coeditor pairs connected by a path.
\item[Talk fraction] Fraction of total revisions made to talk pages.
\item[Mean similarity] Mean Jaccard similarity (by article) with other WikiProjects; a measure of topical complexity.
\item[Mean editors/article] Mean number of editors collaborating on each article in a WikiProject.
\item[Article count] Total number of articles in the WikiProject.
\item[Editor count] Total number of editors working on articles within a WikiProject.
\item[Revision count] Total number of revisions to articles in a WikiProject.
\item[First assessment] Timestamp of first assessment; a measure of how long a WikiProject has been active.
\item[Mean article age] Mean age of articles within a WikiProject.
\end{description}

Our models are summarized in Table \ref{tab:model}.
Min-cut is excluded from all models to avoid collinearity,
as it is highly correlated with degree.
In-degree and out-degree skewness were also highly correlated,
so we only include out-degree skewness
(results are similar with in-degree skewness).
Heavy-tailed variables are log-transformed.
To test the robustness of our results,
we also computed models using cube root instead of logarithmic transformations,
and using only top- and high-importance articles.
The results were qualitatively similar results for all variables,
except for degree-skewness, which had an inconsistent sign across models.

We see that B-efficiency and C-efficiency have very similar models, but that A-efficiency behaves
differently in its dependence on degree skewness and connectivity.
The different behavior of A-efficiency is likely explained by the observation that the A-Class quality is
infrequently used in practice, meaning that the quality level is usually achieved when an article is rated
as a good or featured article, which involves a different consensus process than lower ratings.

@@

Without additional constraints,
the above model is missing
several key properties of real-world collaborations.
In designing our agent-based model, we paid attention to the
following properties.
\begin{description}
\item[Limited concern] Agents are concerned only with a subset of the entire
state when making decisions and determining preferences.
(On Wikipedia, editors typically interact with a small subset of the articles.)
\item[Concern-based network] Agents interact with other agents who share a
common concern over some subset of the state.
(On Wikipedia, editors interact with others who share interests in the same articles.)
\item[Unknown objective] Agents rank states in order of preference,
but do not have access to the objective function.
(On Wikipedia, there is no ground truth measure of quality.)
\item[Single source of truth] At any given time, the system is in a single state
and agent preferences are based on local modifications to that state.
(At any point in time, there is only one current version of Wikipedia.)
\end{description}

>>>>

\begin{abstract}
The internet has enabled collaborations at a scale never before possible,
but the best practices for organizing such large collaborations are still not clear.
Wikipedia is a visible and successful example of such a collaboration which might offer
insight into what makes large-scale, decentralized collaborations successful.
We analyze the relationship between the structural properties of WikiProject coeditor networks
and the performance and efficiency of those projects.
We confirm the existence of an overall performance-efficiency trade-off,
while observing that some projects are higher than others in both performance
and efficiency,
suggesting the existence factors correlating positively with both.
Namely, we find an association between low-degree coeditor networks
and both high performance and high efficiency.
We also confirm results seen in previous numerical and small-scale lab studies:
higher performance with less skewed node distributions,
and higher performance with shorter path lengths.
We use agent-based models to explore possible mechanisms for
degree-dependent performance and efficiency.
We present a novel local-majority learning strategy designed to satisfy properties
of real-world collaborations.
The local-majority strategy as well as a localized conformity-based strategy
both show degree-dependent performance and efficiency,
but in opposite directions,
suggesting that these factors depend on both network structure and learning strategy.
Our results suggest 
possible benefits to decentralized collaborations made of smaller,
more tightly-knit teams,
and that these benefits may be modulated by the particular learning strategies
in use.
\end{abstract}

@@

We look at WikiProjects on the English-language Wikipedia.
WikiProjects are collections of thematically related articles,
each with their own standards and norms.
When measuring the quality of collaborative projects,
there are at least two distinct measures to consider.
The first measure is short-term:
how effective a unit of work is at improving
the collaboration's output,
which we call {\em efficiency}.
The other measure is long-term:
the highest quality typically reached by an output,
which we call {\em performance}.
These two terms are often used interchangeably,
but we find it fruitful to distinguish between the two.
We find that Wikipedia exhibits an overall trade-off between performance and efficiency.
However, some WikiProjects surpass others in both efficiency and performance,
suggesting the existence of factors that correlate positively with both.

Our study focuses on the coeditor networks of each WikiProject:
which editors have edited at least one article in common?
These relationships represent the possible flow of information.
We focus specifically on mean degree, degree skewness, and path length.
High-degree editors have more collaborators,
which can increase diversity and access to information at the possible
expense of higher coordination costs
\cite{hong_groups_2004,golub_naive_2010}.
Highly skewed degree distributions can amplify the biases of high-degree
editors while reducing the need for explicit coordination
\cite{kearns_experiments_2012}.
Networks with shorter path lengths allow information to travel more quickly
at the possible expense of less local diversity
\cite{mason_propagation_2008,barkoczi_social_2016}.

In addition to our empirical study,
we use agent-based modeling to examine the consequences of specific
assumptions on networked collaboration.
We model individual behavior using a {\em social learning strategy} that
assumes agents 1. can only access a fraction of the model's state,
2. interact with others who share their concerns,
and 3. integrate their preferences into a single state.
Our model is the first we are aware of to incorporate these assumptions,
which are present across many real-world collaborations,
including Wikipedia.

Our main findings are:
\begin{itemize}
\item Despite an overall performance/efficiency trade-off,
WikiProjects with low-degree coeditor networks tend to have both higher performance and higher efficiency;
\item Short paths are associated with higher performance, consistent with a conformity-based learning strategy;
\item Structural inequality, as measured by degree skewness, is associated with lower performance;
\item Our agent-based model
shows that the efficiency and performance of collaborations can
depend on network degree,
and that the direction of that dependence varies with social learning strategy.
\end{itemize}

Our findings shed light on the importance of network structure for successful
collaboration.
These findings might be informative for future interventions that
recommend tasks based on how they will
influence network structure,
or for interventions that seek to encourage behaviors
complementary to existing network structure.

\section{Background and Related Work}
\label{sec:background}
The present paper investigates the relationship between social networks
and collaboration outcomes.
This connection has been explored by a number of theoretical,
numerical, and small-scale lab studies in the field of
{\em social learning}.
We contribute to this literature with a large-scale, empirical field study.
In much of the existing literature,
degree distribution correlates with outcome measures.
But aside from the naive Bayes case, it is unknown whether
the correlation is explained best by degree or by another structural property, such as characteristic path length.
In the empirical networks we study,
unlike artificial networks,
the structural properties vary independently,
making it easier to isolate individual
network properties that correlate with outcome variables.

{\bf Social Learning.} In {\em networked social learning},
agents are represented by nodes on a network
and can interact only with their neighbors.
Social learning tasks can be divided into cases where agents have {\em generated signals}
(independently noisy estimates of a true value)
and those where agents have {\em interpreted signals}
(solutions based on different selections of available data)
\cite{hong_interpreted_2009}.
The behavior of individual agents is described by their
{\em social learning strategy}.
For generated signals,
a naive Bayesian approach converges to the truth
when all agents have the same degree,
while the speed of convergence depends on the {\em spectral gap}
between the two largest eigenvalues of the network's interaction matrix
\cite{degroot_reaching_1974,golub_naive_2010}.
Complex social learning tasks can also be modeled as the problem
of maximizing an objective function with many local maxima,
referred to as a {\em rugged landscape}
\cite{lazer_network_2007,mason_propagation_2008,mason_collaborative_2012,grim_scientific_2013,barkoczi_social_2016}.
Numerical simulations have shown that efficient networks
(those with short paths between nodes)
can result in faster convergence at the cost of a less optimal solution,
due to less time for exploration
\cite{mason_propagation_2008,grim_scientific_2013}.
However, when conformity-based social learning strategies are used,
efficient networks can sometimes find more optimal solutions than
inefficient ones \cite{barkoczi_social_2016}.
Using an agent-based model, Hong and Page \shortcite{hong_groups_2004} found that
diverse groups can outperform groups composed of the best individual
problem-solvers.

{\bf Lab experiments.} Lab-based experiments on networked collaboration
suggest a complex interaction between network topology and other factors.
While groups of networked human subjects perform very well on
difficult graph-coloring tasks, the best performing network architectures
(e.g., fully-connected vs. small-world) vary
from task to task \cite{kearns_experiments_2012}.
The same studies found that while human subjects tend to perform well on
many networks, they perform worst on self-organized
networks, possibly due to higher structural inequality
(degree skewness).
Similarly, some network topologies are able to reach faster decisions in the
presence of more information, while others show the opposite effect
\cite{kearns_experimental_2006}.
Based on lab experiments, Fowler and Christakis \shortcite{fowler_cooperative_2010}
suggest that individual decisions towards altruism are conditional on their
neighbor's behavior and ``contagious'' up to three degrees away.
Later experiments by Suri and Watts \shortcite{suri_cooperation_2011} confirmed the
existence of conditional altruism,
but concluded that altruism
influences only first-degree neighbors.

{\bf Digital Communities.}
Research on digital communities has also examined the role of 
diversity and inequality
in collaborative work and decision-making.
In sociology, research has focused on the relationship between
network structure and social capital.
Powerful individuals are often ``brokers''
who act as exclusive intermediaries between disconnected portions of the
social network \cite{silverman_patronage_1965}.
Similarly, successful innovation in organizations often occurs in ``structural
holes'' between groups \cite{granovetter_strength_1973}.

@@

Across the broad range of work discussed above,
a few key themes emerge.
Both the efficiency and the performance of a collaboration are important
considerations and vary depending on both network structure and type of task
\cite{kearns_experiments_2012}.
While generated signal models of social learning predict no relationship
between the two
\cite{golub_naive_2010},
contagion-style innovation models predict a trade-off
\cite{mason_collaborative_2012,barkoczi_social_2016}.
Such a trade-off has been observed in simulations and lab experiments on
collaboration
\cite{kearns_experiments_2012,grim_scientific_2013}.

@@

Many articles on Wikipedia belong to one or more WikiProjects.
WikiProjects are groups of thematically-related articles
(e.g., articles related to Philosophy).
Information about an article's associated WikiProjects
can be viewed on that
article's talk page (Figure \ref{fig:knitting}).
Each WikiProject has its own page and talk page,
containing information about conventions within the project
as well as discussions about individual articles.
WikiProjects are thus distinct communities, with distinct norms and processes.
These communities are the fundamental units of analysis in this paper.

@@

In addition to the above assessment classes, articles on Wikipedia can be tagged as
``good article'' (GA) or ``featured article'' (FA) quality.
FA and GA determinations are made using a Wikipedia-wide consensus,
independently of WikiProject-based evaluations.
FA articles are ``the best articles Wikipedia has to offer''
\cite{wikipedia_contributors_wikipedia:featured_2018}.
GA articles meet ``a core set of editorial standards`` but are ``not featured article quality''
\cite{wikipedia_contributors_wikipedia:good_2017}.
When an article is assigned GA or FA status,
WikiProject quality assessments are often updated to reflect that status.
For example, the article {\em Mewtwo} was assessed as GA status on October 5,
2009 and shortly afterwards
its quality assessment was changed from B to GA within both
{\em WikiProject Pok\'emon} and
{\em WikiProject Video Games}.
This example also illustrates a quirk of conventions on Wikipedia:
very often, articles pass to GA or FA directly from B, skipping A.
The majority of WikiProjects rarely use the A class quality assessment.

@@

For a WikiProject, efficiency quantifies how much participants can raise the
assessed quality of an article for a fixed amount of work.
We measure work by the number of revisions made.
Quality assessments are made through consensus of the project participants themselves.
Different projects can have different standards and practices for assessing article quality,
so the efficiency is not a measure of how quickly some objective measure of quality improves,
but rather of how quickly the project participants can reach consensus on the improvements that
need to be made and make those improvements.
Because our definition relies on assessment transitions,
we must define efficiency variables for
each of the project-level quality assessments: A, B, and C.
For a particular grade $G$,
we desire our definition of efficiency to meet the following conditions:
\begin{itemize}
\item{Strictly increasing in the number of articles reaching grade $G$ (with revision count fixed);}
\item{Strictly decreasing in the number of revisions (with transition count fixed);}
\item{Independent of WikiProject size: not affected by adding an article having the same efficiency.}
\end{itemize}

We now define an efficiency measure which meets the above criteria.
Let $T(W,G)$ be the set of article assessment transitions from below grade $G$
to grade $G$ or higher in project $W$.
Let $N(W,G)$ be the number of articles in project $W$ which ever transition
from below grade $G$ to grade $G$ (or higher).
Given a transition $t$,
let $r(t)$ be the number of revisions to the article
since its previous grade transition,
and let $g(t)$ be the number of grade levels crossed bt $t$.
We quantify the efficiency $E(W,G)$ as the inverse of the mean number of revisions
per transition:
\beq
E(W,G)
&=&
\left[
\frac{1}{
N(W,G)
}
\sum_{t \in T(W,G)} \frac{r(t)}{g(t)}
\right]^{-1},
\eeq
where the $g(t)$ term accounts for assessments that raise article quality by
several grades by
dividing the revisions evenly between all grade levels achieved.

For performance, we wish to quantify how good articles tend to be when they reach a stable state.
Measuring performance is difficult for two reasons:
there is no objective measure of article quality available,
and articles are always changing, making it difficult to know which articles should be considered
complete or stable.
We use an extremely simple performance measure that gives surprisingly consistent results.
In addition to per-project quality assessment, articles can be given ``featured article'' or
``good article'' status.
The criteria for these statuses are consistent across all of Wikipedia,
and any editor can participate in the discussion and decision to award good or featured
status.
In other words, the good and featured statuses are less subjective than per-project assessments.

@@

We would like to determine how the social network structure of
Wikipedia---the pattern of who interacts with whom---relates to
efficiency and performance.
There are several types of interactions we could focus on,
including:
coediting, user talk messages, and talk page replies.
We choose to focus on coediting: when two editors have made changes to the
same article or talk page.
While editors can communicate directly through user talk messages,
the number of such messages is small compared to the number of edits to article
and talk pages.
We also could have considered direct replies between editors on article talk
pages, but these replies are typically seen (and intended to be seen)
by everyone reading the talk page,
and are part of larger conversations.
When an editor views a page,
they are potentially viewing content from and interactions
between all editors who came before them,
motivating our choice to focus on the social network structure of
coeditors.

The {\em coeditor network} of a WikiProject consists of nodes representing editors
and edges connecting pairs of editors who have edited the same article.
The edges are directed, with the direction representing
{\em plausible information flow};
an edge from Alice to Bob exists if Alice edited an article and then Bob edited the same article at
a later time. Note that edges can exist in both directions. 
We make the simplifying assumption of unit weight for all edges.
We focus on three structural properties:
degree, characteristic path length, and min-cut.
Degree and characteristic path length have been shown to correlate with
performance and efficiency in some social learning settings
\cite{golub_naive_2010,mason_propagation_2008,grim_scientific_2013},
while min-cut can be interpreted as a measure of decentralization,
common feature of peer-produced communities such as Wikipedia
\cite{benkler_wealth_2006}.

The degree distribution is the simplest network property we analyze.
The in-degree (out-degree) of a node is the number of edges to (from) that node.
Taking the average of either in-degree or out-degree gives the same value:
the {\em mean degree} of the network.
In our context, the mean degree represents, on average,
how many other editors each editor has collaborated with.
We also consider the {\em skewness} of the in-degree and out-degree distributions.
A large positive degree skewness for a WikiProject coeditor network
implies that a small number of editors have a very large number of collaborators,
while a small positive value implies that the editors having the most collaborators
don't have many more than a typical editor.

@@

Our final network measure quantifies the connectivity of a project's coeditor
network using min-cut size.
The minimum $st$-cut between nodes $s$ and $t$ is the set of edges that must beremoved
for no path exists from $s$ to $t$.
The minimum cut (min-cut) of a graph is the smallest minimum $st$-cut over all node pairs $st$. 
The size of the graph min-cut quantifies the connectivity of a graph,
but only incorporates information about edges lying on paths crossing the min-cut.
Instead, we use the mean size of all minimum $st$-cuts, which we refer to as the
{\em mean min-cut}.
This measure quantifies the number of redundant paths information can take through the network.
Networks with higher redundancy are more resilient to errors on one path \cite{albert_error_2000}
and allow innovations to propagate through complex contagion,
in which innovations are only adopted after multiple exposures \cite{centola_complex_2007}.

The mean path and min-cut are computationally intensive,
requiring distance and minimum $st$-cut calculations for all node pairs.
For larger projects, these calculations are impractical and we thus employed
sampling to determine mean path length and mean min-cut.
For mean path length, source nodes were sampled, and path length was calculated to all destination nodes
from each of these.
For min-cut, node pairs were sampled.
In both cases, stratification was used to ensure the same number of nodes were were sampled from each of
12 node degree quantiles.
We estimated the error due to sampling by determining true values for a medium-sized project,
and calculating error as a function of sample-size.
Sample sizes were chosen such that relative error was below 10\%.
Even with sampling, however, it was impractical to calculate these properties for the largest projects,
so we exclude the 183 largest projects from the analysis.

@@

We also find that mean min-cut is highly correlated with degree ($r=0.980$, $p<0.001$),
so we exclude min-cut from regression models to prevent collinearity.
The high correlation between mean degree and min-cut implies that,
in most cases,
the minimum $st$-cut is simply the set of edges from $s$
or the set of edges to $t$.
The rarity of non-trivial min-cuts suggests that WikiProject coeditor
networks have very few central
bottlenecks and are thus highly decentralized.

@@

To study the relationship between network structure, efficiency, and performance,
we model the performance and efficiency of WikiProjects using ordinary least-squares linear regression.
Each WikiProject is a single observation.
The models include each project's coeditor network properties as independent variables.
We also include the following project-level variables to control for confounding factors.
\begin{description}
\item[C-efficiency]
(performance only).
Quantifies how quickly a WikiProject improves articles.
Efficiencies for different grades are highly correlated,
so we include only one.
\item[Connected fraction.]
Fraction of coeditor pairs connected by a path.
\item[Talk fraction.] Fraction of total revisions made to talk pages.
\item[Mean similarity.] Mean Jaccard similarity (by article) with other WikiProjects; a measure of topical complexity.
\item[Mean editors/article.] Mean number of editors collaborating on each article in a WikiProject.
\item[Article count.] Total number of articles in the WikiProject.
\item[Editor count.] Total number of editors working on articles within a WikiProject.
\item[Revision count.] Total number of revisions to articles in a WikiProject.
\item[First assessment.] Timestamp of first assessment; a measure of how long a WikiProject has been active.
\item[Mean article age.] Mean age of articles within a WikiProject.
\end{description}

Our models are summarized in Table \ref{tab:model}.
Min-cut is excluded from all models to avoid collinearity,
as it is highly correlated with degree.
In-degree and out-degree skewness were also highly correlated,
so we only include out-degree skewness
(results are similar for in-degree skewness).
Heavy-tailed variables are log-transformed.
To test the robustness of our results,
we also computed models using cube root instead of logarithmic transformations,
and using only top- and high-importance articles.
The results were qualitatively similar results for all variables,
except for degree-skewness, which had an inconsistent sign across models.

We see that B-efficiency and C-efficiency have very similar models, but that A-efficiency behaves
differently in its dependence on degree skewness and connectivity.
The different behavior of A-efficiency is likely explained by the
observation that the A-Class quality is infrequently used in practice.
The A-Class quality level is usually passed when an article reaches
good or featured article status,
which follow deifferent a consensus process from other ratings.

@@

Without additional constraints,
the above model is missing
several key properties of real-world collaborations.
In designing our agent-based model, we paid attention to the
following properties.
\begin{description}
\item[Limited concern.] Agents are concerned only with a subset of the entire
state when making decisions and determining preferences.
(On Wikipedia, editors typically interact with a small subset of the articles.)
\item[Concern-based network.] Agents interact with other agents who share a
common concern over some subset of the state.
(On Wikipedia, editors interact with others who share interests in the same articles.)
\item[Unknown objective.] Agents rank states in order of preference,
but do not have access to the objective function.
(On Wikipedia, there is no ground truth measure of quality.)
\item[Single source of truth.] At any given time, the system is in a single state
and agent preferences are based on local modifications to that state.
(At any point in time, there is only one current version of Wikipedia.)
\end{description}
